---
title: "C8_PML_W4_A"
author: "Amyn"
date: "September 19, 2016"
output: html_document
---

# What to Submit

* A link to a Github repo with your R markdown and compiled HTML file describing your analysis. 
* Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. 
* Submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. 

These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Data & Package Loading

## Package Loading
```{r}
library(caret)      ### Load the CARET Package for modeling
library(ggplot2)    ### Load the GGPLOT2 Package for Data Visualization and Analysis
set.seed(1234)      ### Set seed so that we can ensure that we can recreate our results for testing purposes
library(data.table)
library(quantmod)
library(randomForest)
library(rpart)
library(rpart.plot)
setwd("~/Documents/R Working Directory/Course_8_Practical_Machine_Learning")    ### Set the Working Directory 

```

## Load the Data
The training and test data for this project are available using the following links:
* https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
* https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
```{r}

url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
f <- file.path(getwd(), "pml-training.csv")
 download.file(url, f)  ### Commenting out the download to simplify testing
train<- read.csv(f, na.strings = c("NA","#DIV/0!", ""))

url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
f <- file.path(getwd(), "pml-testing.csv")
 download.file(url, f) ### Commenting out the download to simplify the testing
validation<- read.csv(f, na.strings = c("NA","#DIV/0!", ""))
```

# Data Cleansing + Transformation
We will remove all of the columns that nothing by NAs and we will also eliminate the first 7 columns as they don't provide quantitative data for the model
```{r}
#dim(train)
#dim(validation)
train<-train[,colSums(is.na(train)) == 0]   ### Remove all columns that have nothing but NAs - there column sum will not be 0
validation <-validation[,colSums(is.na(validation)) == 0]     ### Remove all columns that have nothing but NAs - there column sum will not be 0
train <- train[,-c(1:7)]    ### Remove the first 7 columns of qualitative data as they are not required for our model
validation <-validation[,-c(1:7)]       ### Remove the first 7 columns of qualitative data as they are not required for our model
#dim(train)
#dim(validation)

```

We will now  partition the training set for cross validation testing activitie
```{r}

inTrain <-createDataPartition(y=train$classe, p = 0.75, list = FALSE)
subTrain <- train[inTrain,] ### Create our Training Set
subTest <-train[-inTrain,]  ### Create our Testing Set
#dim(subTrain)
#dim(subTest)

```


# Modeling
We will use the RandomForest model to simulate our results first.

```{r}

mod_rf <- train(classe ~ ., data = subTrain, method = "rf", trControl = trainControl(method="cv"), number = 3)
mod_rf$finalModel
```
We will also look to see what are the most and least important variables for our model
```{r}
varImpObj <- varImp(mod_rf)
plot(varImpObj, main = "Importance Variables", top = (ncol(subTrain)-1))
```
So we can see that the ROLL_BELT, YAW_BELT & MAGNET_DUBMBELL_Z are the most important, while the GYROS_ARM_Z, GYROS_FOREARM_X and GYROS_DUMBELL_Z are the least imporant in our prediction model.

We will then do the Prediction and review the accuracy and cross correlation with our test data set
```{r}
pred_rf_T <- predict(mod_rf, subTest)
confusionMatrix(pred_rf_T,subTest$classe)
```
From the test we can see that the RandomForest Model works very well our test data set with an Accuracy of 99.4% and a Confidence Interval of 99.1% to 99.6%.  

# Final Validation & Output
We will now do a final validation of the model using the validation data set and output the results to a set of final for confirmation
```{r}
pred_rf_V <-predict(mod_rf,validation)
pred_rf_V

answers <- as.vector(pred_rf_V)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}
pml_write_files(answers)
```

# Acknowledgements

* Data from this project was sourced from http://groupware.les.inf.puc-rio.br/har



